---
title: "Validating `r year` State IO Model"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm
editor_options: 
  chunk_output_type: console
---

This document presents validation results of `r year` summary-level state IO model.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      results = "asis")
logging::basicConfig(level = 40)
```

```{r report-function}
reportValidationResult <- function(failures) {
  if (any(nrow(failures) == 0, length(failures) == 0)) {
    cat("There are no failures.\n\n")
  } else if (any(nrow(failures) == 1, length(failures) == 1)) {
    cat("There is", nrow(failures), "failure, and it is")
    knitr::kable(failures, "simple")
  } else {
    cat("There are", nrow(failures), "failures, and they are")
    knitr::kable(failures, "simple")
  }
}
```

### Prepare data
#### 0. Load state and two-region IO data.
```{r load-data}
devtools::load_all()
# Define basic IO variabls
commodities <- getVectorOfCodes("Summary", "Commodity")
industries <- getVectorOfCodes("Summary", "Industry")
FD_cols <- getFinalDemandCodes("Summary")
VA_rows <- getVectorOfCodes("Summary", "ValueAdded")
# Load Us Make, Use, and DomesticUse
US_Summary_Make <- getNationalMake("Summary", year)
US_Summary_Use <- getNationalUse("Summary", year)
US_Summary_DomesticUse <- generateUSDomesticUse("Summary", year)
# Load state Make, industry and commodity output
State_Summary_Make_ls <- loadStateIODataFile(paste0("State_Summary_Make_", year))
states <- names(State_Summary_Make_ls)
# Load state total Use and domestic Use
State_Summary_Use_ls <- loadStateIODataFile(paste0("State_Summary_Use_", year))
State_Summary_DomesticUse_ls <- loadStateIODataFile(paste0("State_Summary_DomesticUse_", year))
State_Summary_IndustryOutput_ls <- loadStateIODataFile(paste0("State_Summary_IndustryOutput_", year))
State_Summary_CommodityOutput_ls <- loadStateIODataFile(paste0("State_Summary_CommodityOutput_", year))
# Load two-region tables
TwoRegionMake_ls <- loadStateIODataFile(paste0("TwoRegion_Summary_Make_", year))
TwoRegionCommodityOutput_ls <- loadStateIODataFile(paste0("TwoRegion_Summary_CommodityOutput_", year))
TwoRegionIndustryOutput_ls <- loadStateIODataFile(paste0("TwoRegion_Summary_IndustryOutput_", year))
TwoRegionDomesticUse_ls <- loadStateIODataFile(paste0("TwoRegion_Summary_DomesticUse_", year))
TwoRegionDomesticUsewithTrade_ls <- loadStateIODataFile(paste0("TwoRegion_Summary_DomesticUsewithTrade_", year))
TwoRegionITA_ls <- loadStateIODataFile(paste0("TwoRegion_Summary_InternationalTradeAdjustment_", year))
cat("State and two-region IO data successfully loaded.")
```

### Check state IO tables
#### 1. Check if industry output from state Make and Use are almost equal (<= 0.01).
```{r check1}
failures <- c()
for (state in states[states != "Overseas"]) {
  # Calculate Industry Output (x) from Make and Use
  x_make <- rowSums(State_Summary_Make_ls[[state]])
  x_use <- colSums(State_Summary_Use_ls[[state]][, rownames(US_Summary_Make)])
  # Calculate relative differences in x_make and x_use
  rel_diff <- (x_use - x_make)/x_make
  # Compare rel_diff against 0 with a tolerance of 1E-3
  failures_state <- formatValidationResult(rel_diff, abs_diff = TRUE, tolerance = 0.01)[["Failure"]]
  if (nrow(failures_state) > 0) {
    failures_state$State <- state
  }
  colnames(failures_state)[1:3] <- c("Industry", "Relative Diff", "Validation")
  failures <- rbind(failures, failures_state)
}
cat("\n")
reportValidationResult(failures)
```

#### 2. Check commodity output
#### 2.1. Check if commodity output from state Make and Use are almost equal (relative difference <= 0.01).
```{r check2.1}
failures <- c()
for (state in states[states != "Overseas"]) {
  # Calculate Commodity Output (q) from Make and Use
  q_make <- colSums(State_Summary_Make_ls[[state]])
  q_use <- rowSums(State_Summary_Use_ls[[state]][rownames(US_Summary_Use), colnames(US_Summary_Use)])
  # Calculate relative differences in q_make and q_use
  rel_diff <- (q_use - q_make)/q_make
  # Compare rel_diff against 0 with a tolerance of 1E-3
  failures_state <- formatValidationResult(rel_diff, abs_diff = TRUE, tolerance = 0.01)[["Failure"]]
  if (nrow(failures_state) > 0) {
    failures_state$State <- state
  }
  colnames(failures_state)[1:3] <- c("Commodity", "Relative Diff", "Validation")
  failures <- rbind(failures, failures_state)
}
cat("\n")
reportValidationResult(failures)
```

#### 2.2. Check if commodity output from state Use and Domestic Use are almost equal (relative difference <= 0.01).
```{r check2.2}
failures <- c()
for (state in states[states != "Overseas"]) {
  # Calculate Commodity Output (q) from Make and Use
  q_use <- rowSums(State_Summary_Use_ls[[state]][rownames(US_Summary_Use), colnames(US_Summary_Use)]) # doesn't include F051
  q_domuse <- rowSums(State_Summary_DomesticUse_ls[[state]][rownames(US_Summary_Use), ]) # includs F051
  # Calculate relative differences in q_use and q_domuse
  rel_diff <- (q_domuse - q_use)/q_use
  # Compare rel_diff against 0 with a tolerance of 1E-3
  failures_state <- formatValidationResult(rel_diff, abs_diff = TRUE, tolerance = 0.01)[["Failure"]]
  if (nrow(failures_state) > 0) {
    failures_state$State <- state
  }
  colnames(failures_state)[1:3] <- c("Commodity", "Relative Diff", "Validation")
  failures <- rbind(failures, failures_state)
}
cat("\n")
reportValidationResult(failures)
```

### Compare state to US IO tables (negativity in the same cell & state sum == US totals)
#### 3. Sum of each cell across all state Make tables must almost equal (<= 0.001) the same cell in US Make table.
```{r check3}
df0 <- US_Summary_Make
df1 <- Reduce("+", State_Summary_Make_ls)
rownames(df1) <- gsub(".*\\.", "", rownames(df1))
df1 <- df1[rownames(df0), colnames(df0)]
# Compare aggregated state Make against US Make
failures <- formatValidationResult(df1 - df0, abs_diff = TRUE, tolerance = 1E-3)[["Failure"]]
reportValidationResult(failures)
```

#### 4. There should not be any negative values in state Make table, unless they are negative in US Make table.
Note: only exception being Overseas, which isn't used for further calculations, and if the same cell in US Make table is also negative.
```{r check4}
US_failures <- formatValidationResult(US_Summary_Make - abs(US_Summary_Make),
                                      abs_diff = TRUE, tolerance = 0)[["Failure"]]
colnames(US_failures) <- c("Industry", "Commodity")
# Validate Make table and extract failures for each state
failures <- data.frame()
for (state in states[states != "Overseas"]) {
  df <- as.data.frame(State_Summary_Make_ls[[state]])
  # Check if there is zero in state Make table
  failures_state <- formatValidationResult(df - abs(df), abs_diff = TRUE,
                                           tolerance = 0)[["Failure"]]
  colnames(failures_state) <- c("Industry", "Commodity")
  # If failure is in US Make table, remove it from state failures
  if (nrow(failures_state) > 0) {
    failures_state$index <- paste(sub(".*\\.", "", failures_state$Industry),
                                  failures_state$Commodity)
    US_state_diff <- setdiff(failures_state$index,
                             paste(US_failures$Industry, US_failures$Commodity))
    failures_state <- failures_state[failures_state$index %in% US_state_diff, ]
  }
  failures_state$index <- rownames(failures_state) <- NULL
  failures <- rbind(failures, failures_state)
}
reportValidationResult(failures)
```

#### 5. Sum of each industry's output across all states must almost equal (<= 1E7, or $10 million by industry) the industry output in US Make Table.
The threshold is set to 1E7 because there are differences (within +/- $10 million) between US industry output summed from Make and that summed from Use, comparing sum of state industry output (summed from state Use) to US industry output summed from US Make should account for those inherent differences at the national level.
```{r check5}
df0 <- as.data.frame(rowSums(US_Summary_Make))
df1 <- Reduce("+", State_Summary_IndustryOutput_ls)
rownames(df1) <- gsub(".*\\.", "", rownames(df1))
df1 <- df1[rownames(df0), ]
# Compare aggregated state industry output against US industry output
failures <- formatValidationResult(df1 - df0, abs_diff = TRUE, tolerance = 1E7)[["Failure"]]
colnames(failures) <- c("Industry", "")
reportValidationResult(failures)
```

#### 6. Sum of each commodity's output across all states must almost equal (<= 1E7, or $10 million by commodity) the commodity output in US Make Table.
The threshold is set to 1E7 because there are differences (within +/- $10 million) between US industry output summed from Make and that summed from Use, comparing sum of state industry output (summed from state Use) to US industry output summed from US Make should account for those inherent differences at the national level.
```{r check6}
df0 <- as.data.frame(colSums(US_Summary_Make))
df1 <- Reduce("+", State_Summary_CommodityOutput_ls)
rownames(df1) <- gsub(".*\\.", "", rownames(df1))
df1 <- df1[rownames(df0), ]
# Compare aggregated state commodity output against US commodity output
failures <- formatValidationResult(df1 - df0, abs_diff = TRUE, tolerance = 1E7)[["Failure"]]
colnames(failures) <- c("Commodity", "")
reportValidationResult(failures)
```

#### 7. Sum of each commodity's output across all states must almost equal (<= 1E^7, or $10 million by commodity) commodity output in US Use Table.
Note: even if the threshold is met, track the difference for each commodity. Save result as a type of quality control check.

```{r check7}
df0 <- as.data.frame(rowSums(US_Summary_Use))
df1 <- Reduce("+", State_Summary_CommodityOutput_ls)
rownames(df1) <- gsub(".*\\.", "", rownames(df1))
df1 <- df1[rownames(df0), ]
# Compare aggregated state commodity output against row sum of US domestic Use
failures <- formatValidationResult(df1 - df0, abs_diff = TRUE, tolerance = 1.11E7)[["Failure"]]
colnames(failures) <- c("Commodity", "")
reportValidationResult(failures)
diff <- df1 - df0
colnames(diff) <- "q_state_sum - q_US_use"
knitr::kable(diff, "simple")
```

#### 8. All cells that are zero in US Make table must remain zero in state Make tables. Find zero values in US Make table.
```{r check8}
US_zeros <- formatValidationResult(US_Summary_Make - 0, abs_diff = TRUE,
                                   tolerance = 0)[["Pass"]]
colnames(US_zeros) <- c("Industry", "Commodity")
failures <- data.frame()
for (state in states) {
  # Find zero values in state Make table
  state_zeros <- formatValidationResult(State_Summary_Make_ls[[state]] - 0,
                                        abs_diff = TRUE, tolerance = 0)[["Pass"]]
  colnames(state_zeros)[1:2] <- c("Industry", "Commodity")
  state_zeros$index <- paste(sub(".*\\.", "", state_zeros$Industry), state_zeros$Commodity)
  # Find difference between state_zeros and US_zeros: zeros in US that are not zeros in state
  US_state_diff <- setdiff(paste(US_zeros$Industry, US_zeros$Commodity), state_zeros$index)
  failures_state <- state_zeros[state_zeros$index %in% US_state_diff, ]
  failures_state$index <- rownames(failures_state) <- NULL
  failures <- rbind(failures, failures_state)
}
reportValidationResult(failures)
```

#### 9. Sum of each cell across all state Use tables must almost equal (<= 1E5, or $5 million) the same cell in US Use table. This validates that Total state demand == Total national demand.
Note: failures associated with 'F050 - Imports' are acceptable. Because state imports are not directly derived from US imports, a gap in imports between state sum and national total is reasonable.

```{r check9}
# Create a function to implement the validation
validateStateUseAgainstNationlUse <- function(domestic = FALSE, rel_diff = FALSE) {
  if (domestic) {
    Use_ls <- State_Summary_DomesticUse_ls
    df0 <- US_Summary_DomesticUse
  } else {
    Use_ls <- State_Summary_Use_ls
    df0 <- US_Summary_Use
  }
  df1 <- Reduce("+", Use_ls)
  rownames(df1) <- gsub(".*\\.", "", rownames(df1))
  df1 <- df1[rownames(df0), colnames(df0)]
  # Compare aggregated state Use table against US Use table
  if (rel_diff) {
    failures <- formatValidationResult((df1 - df0)/df0, abs_diff = TRUE, tolerance = 1E-3)[["Failure"]]
    rownames(failures) <- NULL
    colnames(failures) <- c("Commodity", "Industry/Final Demand", "Relative Diff", "Validation")
    # Provide values in df0 and df1 for failures
    if (nrow(failures) > 0) {
      failures[, c("US", "StateSum")] <- 0
      for (i in 1:nrow(failures)) {
        row <- failures[i, "Commodity"]
        col <- failures[i, "Industry/Final Demand"]
        failures[i, "US"] <- df0[row, col]
        failures[i, "StateSum"] <- df1[row, col]
      }
    }
  } else {
    failures <- formatValidationResult(df1 - df0, abs_diff = TRUE, tolerance = 5E6)[["Failure"]]
    rownames(failures) <- NULL
    colnames(failures) <- c("Commodity", "Industry/Final Demand", "Absolute Diff", "Validation")
    # Compare failures to US values
    if (nrow(failures) > 0) {
      failures[, "AbsDiffPortioninNationalTotals"] <- 0
      for (i in 1:nrow(failures)) {
        row <- failures[i, "Commodity"]
        col <- failures[i, "Industry/Final Demand"]
        failures[i, "AbsDiffPortioninNationalTotals"] <- failures[i, "Absolute Diff"]/df0[row, col]
      }
    }
  }
  
  return(failures)
}
StateUseValidationFailures_absdiff <- validateStateUseAgainstNationlUse(domestic = FALSE, rel_diff = FALSE)
StateDomesticUseValidationFailures_absdiff <- validateStateUseAgainstNationlUse(domestic = TRUE, rel_diff = FALSE)
StateUseValidationFailures_reldiff <- validateStateUseAgainstNationlUse(domestic = FALSE, rel_diff = TRUE)
StateDomesticUseValidationFailures_reldiff <- validateStateUseAgainstNationlUse(domestic = TRUE, rel_diff = TRUE)

# State Use tables
cat("##### 9.1 State Use tables (checking absolute differences)\n")
reportValidationResult(StateUseValidationFailures_absdiff)
# State Domestic Use tables
cat("##### 9.2 State Domestic Use tables (checking absolute differences)\n")
reportValidationResult(StateDomesticUseValidationFailures_absdiff)
# State Use tables
cat("##### 9.3 State Use tables (checking relative differences)\n")
reportValidationResult(StateUseValidationFailures_reldiff)
# State Domestic Use tables
cat("##### 9.4 State Domestic Use tables (checking relative differences)\n")
reportValidationResult(StateDomesticUseValidationFailures_reldiff)
```

### Check two-region model results
#### 10. If SoI commodity output == 0, SoI2SoI ICF ratio == 0
```{r check10}
failures <- c()
for (state in states[states != "Overseas"]) {
  # Find zero values in SoI commodity output
  CO_zeros <- formatValidationResult(State_Summary_CommodityOutput_ls[[state]] - 0,
                                     abs_diff = TRUE, tolerance = 0)[["Pass"]]
  colnames(CO_zeros) <- c("Commodity", "")
  # Find zero values in SoI2SoI ICF ratio
  ICF <- generateDomestic2RegionICFs(state, year, ioschema = 2012, iolevel = "Summary",
                                     ICF_sensitivity_analysis = FALSE, adjust_by = 0)
  SoI2SoI_ICF_zeros <- ICF[ICF$SoI2SoI == 0, 1]
  # Find difference between SoI2SoI_ICF_zeros and CO_zeros:
  # zeros in CO that are not zeros in SoI2SoI_ICF
  diff <- setdiff(CO_zeros$Commodity, SoI2SoI_ICF_zeros)
  failures_state <- SoI2SoI_ICF_zeros[SoI2SoI_ICF_zeros %in% diff]
  failures <- c(failures, failures_state)
  print(state)
}
cat("\n")
reportValidationResult(failures)
```

#### 11. SoI and RoUS interregional exports >= 0, interregional imports >= 0
```{r check11}
failures <- data.frame()
for (state in states[states != "Overseas"]) {
  # Prepare domestic 2-region Use tables
  TwoRegionTable_state <- TwoRegionDomesticUsewithTrade_ls[[state]]
  df <- cbind(TwoRegionTable_state[["SoI2SoI"]][, c("InterregionalImports", "InterregionalExports")],
              TwoRegionTable_state[["RoUS2RoUS"]][, c("InterregionalImports", "InterregionalExports")])
  df <- as.data.frame(sapply(df, round, 1))
  rownames(df) <- rownames(TwoRegionTable_state[["SoI2SoI"]])
  colnames(df) <- paste(rep(paste(state, c("SoI2SoI", "RoUS2RoUS"), sep = "_"), each = 2),
                        colnames(df), sep = "$")
  # Compare SoI and RoUS interregional exports and imports against 0
  failures_state <- formatValidationResult(abs(df) - df, abs_diff = FALSE, tolerance = 0)[["Failure"]]
  failures <- rbind(failures, failures_state)
}
reportValidationResult(failures)
```

#### 12. SoI net exports + RoUS net exports == 0
```{r check12}
failures <- data.frame()
for (state in states[states != "Overseas"]) {
  # Prepare domestic 2-region Use tables
  TwoRegionTable_state <- TwoRegionDomesticUsewithTrade_ls[[state]]
  df <- TwoRegionTable_state[["SoI2SoI"]][, "NetExports", drop = FALSE] +
    TwoRegionTable_state[["RoUS2RoUS"]][, "NetExports", drop = FALSE]
  df <- as.data.frame(sapply(df, round, 1))
  rownames(df) <- rownames(TwoRegionTable_state[["SoI2SoI"]])
  # Compare SoI net exports + RoUS net exports against 0
  failures_state <- formatValidationResult(df, abs_diff = TRUE, tolerance = 0)[["Failure"]]
  failures <- rbind(failures, failures_state)
}
reportValidationResult(failures)
```

#### 13. Check row sum of SoI2SoI <= state commodity supply. Row sum of RoUS2RoUS <= RoUS commodity supply.
```{r check13}
failures <- data.frame()
for (state in states[states != "Overseas"]) {
  # Prepare state commodity supply
  df0 <- cbind.data.frame(State_Summary_CommodityOutput_ls[[state]][, "Output", drop = FALSE],
                          colSums(US_Summary_Make))
  colnames(df0) <- c("StateCommOutput", "USCommOutput")
  df0$RoUSCommOutput <- df0$USCommOutput - df0$StateCommOutput
  df0$USCommOutput <- NULL
  # Prepare domestic 2-region Use tables
  TwoRegionTable_state <- TwoRegionDomesticUsewithTrade_ls[[state]]
  columns <- c(getVectorOfCodes("Summary", "Industry"), getFinalDemandCodes("Summary"), "ExportResidual")
  df1 <- cbind.data.frame(rowSums(TwoRegionTable_state[["SoI2SoI"]][, columns]),
                          rowSums(TwoRegionTable_state[["RoUS2RoUS"]][, columns]))
  colnames(df1) <- c(state, paste0(state, "'s RoUS"))
  # Compare row sum of SoI2SoI against state commodity supply
  failures_state <- formatValidationResult(df1 - df0, abs_diff = FALSE,
                                           tolerance = 1E7)[["Failure"]]
  failures <- rbind(failures, failures_state)
}
reportValidationResult(failures)
```

#### 14. Value in SoI2SoI and RoUS2RoUS can be negative only when the same cell is negative in national Use table
```{r check14}
# Find negative values in US Use table
US_negatives <- formatValidationResult(abs(US_Summary_DomesticUse) - US_Summary_DomesticUse,
                                       abs_diff = FALSE, tolerance = 0)[["Failure"]]
colnames(US_negatives) <- c("Commodity", "Industry")
# Validate the position of zero values in state Make tables
failures <- data.frame()
for (state in states[states != "Overseas"]) {
  # Prepare domestic 2-region Use tables
  TwoRegionTable_state <- TwoRegionDomesticUsewithTrade_ls[[state]]
  columns <- c(getVectorOfCodes("Summary", "Industry"), getFinalDemandCodes("Summary"))
  # Find negative values in SoI2SoI Use table
  df_SoI <- TwoRegionTable_state[["SoI2SoI"]][, columns]
  SoI_negatives <- formatValidationResult(abs(df_SoI) - df_SoI, abs_diff = FALSE,
                                          tolerance = 0)[["Failure"]]
  colnames(SoI_negatives) <- c("Commodity", "Industry")
  SoI_negatives$index <- paste(SoI_negatives$Commodity, SoI_negatives$Industry)
  SoI_negatives$table <- paste(state, "SoI2SoI Use")
  # Find difference between SoI_negatives and US_negatives:
  # negatives in SoI2SoI that are not negatives in US Use
  US_SoI_diff <- setdiff(SoI_negatives$index,
                         paste(US_negatives$Commodity, US_negatives$Industry))
  # Find negative values in RoUS2RoUS Use table
  df_RoUS <- TwoRegionTable_state[["RoUS2RoUS"]][, columns]
  RoUS_negatives <- formatValidationResult(abs(df_RoUS) - df_RoUS, abs_diff = FALSE,
                                           tolerance = 0)[["Failure"]]
  colnames(RoUS_negatives) <- c("Commodity", "Industry")
  RoUS_negatives$index <- paste(RoUS_negatives$Commodity, RoUS_negatives$Industry)
  RoUS_negatives$table <- paste(state, "RoUS2RoUS Use")
  # Find difference between RoUS_negatives and US_negatives:
  # negatives in RoUS2RoUS that are not negatives in US Use
  US_RoUS_diff <- setdiff(RoUS_negatives$index,
                          paste(US_negatives$Commodity, US_negatives$Industry))
  # Compile failures
  failures_state <- rbind(SoI_negatives[SoI_negatives$index %in% US_SoI_diff, ],
                          RoUS_negatives[RoUS_negatives$index %in% US_RoUS_diff, ])
  failures_state$index <- rownames(failures_state) <- NULL
  failures <- rbind(failures, failures_state)
}
reportValidationResult(failures)
```

#### 15. SoI interregional imports == RoUS interregional exports, or difference <= 0.001
```{r check15}
failures <- data.frame()
for (state in states[states != "Overseas"]) {
  # Prepare domestic 2-region Use tables
  TwoRegionTable_state <- TwoRegionDomesticUsewithTrade_ls[[state]]
  # Prepare df0 and df1
  df0 <- TwoRegionTable_state[["SoI2SoI"]][, "InterregionalImports", drop = FALSE]
  df1 <- TwoRegionTable_state[["RoUS2RoUS"]][, "InterregionalExports", drop = FALSE]
  # Compare SoI interregional imports against RoUS interregional exports
  failures_state <- formatValidationResult(df0 - df1, abs_diff = FALSE,
                                           tolerance = 1E-3)[["Failure"]]
  if (nrow(failures_state) > 0) {
    failures_state$State <- state
  }
  failures <- rbind(failures, failures_state)
}
reportValidationResult(failures)
```

#### 16.1. Total state commodity supply == state demand by intermediate consumption, plus final demand (except imports and international trade adjustment) + Interregional Exports + Export Residual. Difference must be <= 0.001.
```{r check16.1}
failures <- data.frame()
for (state in states[states != "Overseas"]) {
  # Prepare domestic 2-region Use tables
  TwoRegionTable_state <- TwoRegionDomesticUsewithTrade_ls[[state]]
  columns <- c(getVectorOfCodes("Summary", "Industry"),
               setdiff(getFinalDemandCodes("Summary"),
                       getVectorOfCodes("Summary", "Import")),
               "F051", "InterregionalExports", "ExportResidual")
  # Prepare df0 and df1
  df0 <- State_Summary_CommodityOutput_ls[[state]][, "Output", drop = FALSE]
  df1 <- as.data.frame(rowSums(TwoRegionTable_state[["SoI2SoI"]][, columns]))
  # Compare state commodity supply against state demand by intermediate consumption,
  # plus final demand (except imports) + Interregional Exports
  failures_state <- formatValidationResult(df0 - df1, abs_diff = TRUE,
                                           tolerance = 1E-3)[["Failure"]]
  if (nrow(failures_state) > 0) {
    failures_state$State <- state
  }
  failures <- rbind(failures, failures_state)
}
reportValidationResult(failures)
```

#### 16.2. Total SoI and RoUS commodity supply (output) == SoI and RoUS demand (domestic intermediate consumption + ITA + Export Residual).
```{r check16.2}
failures <- c()
for (state in states[states != "Overseas"]) {
  # Calculate Commodity Output (q) from Make and Use
  q <- TwoRegionCommodityOutput_ls[[state]]
  q_2r_use <- rowSums(TwoRegionDomesticUse_ls[[state]]) + TwoRegionITA_ls[[state]] + c(TwoRegionDomesticUsewithTrade_ls[[state]][["SoI2SoI"]][, "ExportResidual"], TwoRegionDomesticUsewithTrade_ls[[state]][["RoUS2RoUS"]][, "ExportResidual"])
  # Calculate relative differences in q_make and q_use
  rel_diff <- (q - q_2r_use[names(q)])/q_2r_use[names(q)]
  # Compare rel_diff against 0 with a tolerance of 1E-3
  failures_state <- formatValidationResult(rel_diff, abs_diff = TRUE, tolerance = 0.01)[["Failure"]]
  if (nrow(failures_state) > 0) {
    failures_state$State <- state
  }
  colnames(failures_state)[1:3] <- c("Commodity", "Relative Diff", "Validation")
  failures <- rbind(failures, failures_state)
}
cat("\n")
reportValidationResult(failures)
```

#### 17. Number of negative cells in SoI2SoI, SoI2RoUS, RoUS2SoI and RoUS2RoUS <= Number of negative cells in national Use table
```{r check17}
US_negatives <- formatValidationResult(abs(US_Summary_DomesticUse) - US_Summary_DomesticUse,
                                       abs_diff = FALSE, tolerance = 0)[["Failure"]]
colnames(US_negatives) <- c("Commodity", "Industry")
failure <- data.frame()
for (state in states[states != "Overseas"]) {
  # Prepare domestic 2-region Use tables
  TwoRegionTable_state <- TwoRegionDomesticUsewithTrade_ls[[state]]
  columns <- c(getVectorOfCodes("Summary", "Industry"), getFinalDemandCodes("Summary"))
  failures_state <- data.frame()
  for (table in names(TwoRegionTable_state)[1:4]) {
    # Find number of negative cells in two-region Use table
    df <- TwoRegionTable_state[[table]][, columns]
    state_negatives <- formatValidationResult(abs(df) - df, abs_diff = FALSE,
                                              tolerance = 0)[["Failure"]]
    colnames(state_negatives) <- c("Commodity", "Industry")
    state_negatives$index <- paste(state_negatives$Commodity, state_negatives$Industry)
    state_negatives$table <- paste(state, table, "Use")
    # Find difference between state_negatives and US_negatives:
    # negatives in RoUS2RoUS that are not negatives in US Use
    US_state_diff <- setdiff(state_negatives$index,
                             paste(US_negatives$Commodity, US_negatives$Industry))
    # Compile failures
    failures_table <- state_negatives[state_negatives$index %in% US_state_diff, ]
    failures_table$index <- NULL
    failures_table <- rbind(failures_state, failures_table)
  }
  failures <- rbind(failures, failures_state)
}
reportValidationResult(failures)
```

#### 18. Non-square model verification. Validate L matrix of two-region model and final demand against SoI and RoUS output.
##### Absolute difference: L*y - output <= 1^6, or $1 million.
##### Relative difference: (L*y - output)/output <= 1^-2, or 1%.
```{r check18, include=FALSE}
#' @param state A text value specifying state of interest.
#' @param year A numeric value between 2007 and 2017 specifying the year of interest.
#' @param ioschema A numeric value of either 2012 or 2007 specifying the io schema year.
#' @param iolevel BEA sector level of detail, can be "Detail", "Summary", or "Sector".
#' @return A list of validation components and result.
validateTwoRegionLagainstOutput <- function(state, year, ioschema, iolevel) {
  startLogging()
  # Define industries and commodities
  industries <- getVectorOfCodes(iolevel, "Industry")
  commodities <- getVectorOfCodes(iolevel, "Commodity")
  ita_column <- ifelse(iolevel == "Detail", "F05100", "F051")
  # Define state abbreviation
  state_abb <- state.abb[which(state.name == state)]
  logging::loginfo("Generating A matrix of SoI Make table ...")
  # SoI Make
  TwoRegionMake <- TwoRegionMake_ls[[state]]
  SoI_Make <- TwoRegionMake[endsWith(rownames(TwoRegionMake), state_abb),
                            endsWith(colnames(TwoRegionMake), state_abb)]
  # SoI commodity output
  # SoI_Commodity_Output <- State_Summary_CommodityOutput_ls[[state]]
  TwoRegionCommodityOutput <- TwoRegionCommodityOutput_ls[[state]]
  SoI_Commodity_Output <- TwoRegionCommodityOutput[endsWith(names(TwoRegionCommodityOutput),
                                                            state_abb)]
  # SoI A matrix
  SoI_A <- useeior:::normalizeIOTransactions(SoI_Make, SoI_Commodity_Output)
  # Check column sums of SoI_A
  if (all(abs(colSums(SoI_A) - 1) < 1E-3)) {
    logging::loginfo("FACT CHECK: column sums of A matrix of SoI Make table == 1.")
  } else {
    logging::logwarn("Column sums of A matrix of SoI Make table != 1")
  }

  logging::loginfo("Generating A matrix of RoUS Make table ...")
  # RoUS Make
  RoUS_Make <- TwoRegionMake[endsWith(rownames(TwoRegionMake), "RoUS"),
                             endsWith(colnames(TwoRegionMake), "RoUS")]
  # RoUS commodity output
  RoUS_Commodity_Output <- TwoRegionCommodityOutput[endsWith(names(TwoRegionCommodityOutput),
                                                            "RoUS")]

  # RoUS A matrix
  RoUS_A <- useeior:::normalizeIOTransactions(RoUS_Make, RoUS_Commodity_Output)
  # Check column sums of RoUS_A
  if (all(abs(colSums(RoUS_A) - 1) < 1E-3)) {
    logging::loginfo("FACT CHECK: column sums of A matrix of RoUS Make table == 1.")
  } else {
    logging::logerror("Column sums of A matrix of RoUS Make table != 1")
  }

  # Two-region A matrix
  #                       Commodity                Commodity                  Industry                     Industry
  #                          SoI                     RoUS                        SoI                         RoUS
  #               +-----------------------+-------------------------+---------------------------+----------------------------+
  # Commodity SoI |           0           |            0            |norm(SoI2SoI_Use,  SoI_TIO)|norm(SoI2RoUS_Use, RoUS_TIO)|
  #               +-----------------------+-------------------------+---------------------------+----------------------------+
  # Commodity RoUS|           0           |            0            |norm(RoUS2SoI_Use, SoI_TIO)|norm(RoUS2RoUS_Use,RoUS_TIO)|
  #               +-----------------------+-------------------------+---------------------------+----------------------------+
  # Industry  SoI |norm(SoI_Make, SoI_TCO)|            0            |             0             |             0              |
  #               +-----------------------+-------------------------+---------------------------+----------------------------+
  # Industry  RoUS|           0           |norm(RoUS_Make, RoUS_TCO)|             0             |             0              |
  #               +-----------------------+-------------------------+---------------------------+----------------------------+
  #                           Total column sum must equal 1                 Total column sum can but shouldn't equal 1

  logging::loginfo("Generating two-region Domestic Use tables ...")
  ls <- TwoRegionDomesticUsewithTrade_ls[[state]]
  TwoRegionIndustryOutput <- TwoRegionIndustryOutput_ls[[state]]
  SoI_Industry_Output <- TwoRegionIndustryOutput[endsWith(names(TwoRegionIndustryOutput),
                                                          state_abb)]
  RoUS_Industry_Output <- TwoRegionIndustryOutput[endsWith(names(TwoRegionIndustryOutput),
                                                          "RoUS")]
  # If industry/comm output == 0, it's not viable to generate A matrix, hence set it to 1.
  SoI_Industry_Output[SoI_Industry_Output == 0] <- 1

  logging::loginfo("Generating A matrix of SoI2SoI Domestic Use table ...")
  SoI2SoI_A <- useeior:::normalizeIOTransactions(ls[["SoI2SoI"]][, industries],
                                                 SoI_Industry_Output)

  logging::loginfo("Generating A matrix of RoUS2SoI Domestic Use table ...")
  RoUS2SoI_A <- useeior:::normalizeIOTransactions(ls[["RoUS2SoI"]][, industries],
                                                  SoI_Industry_Output)

  logging::loginfo("Generating A matrix of SoI2RoUS Domestic Use table ...")
  SoI2RoUS_A <- useeior:::normalizeIOTransactions(ls[["SoI2RoUS"]][, industries],
                                                  RoUS_Industry_Output)

  logging::loginfo("Generating A matrix of RoUS2RoUS Domestic Use table ...")
  RoUS2RoUS_A <- useeior:::normalizeIOTransactions(ls[["RoUS2RoUS"]][, industries],
                                                   RoUS_Industry_Output)

  logging::loginfo("Assembling the complete A matrix ...")
  # Assemble A matrix
  A_top <- cbind(diag(rep(0, length(commodities)*2)),
                 cbind(rbind(SoI2SoI_A, RoUS2SoI_A),
                       rbind(SoI2RoUS_A, RoUS2RoUS_A)))
  colnames(A_top) <- c(1:ncol(A_top))
  A_btm <- cbind(as.matrix(Matrix::bdiag(list(as.matrix(SoI_A),
                                              as.matrix(RoUS_A)))),
                 diag(rep(0, length(industries)*2)))
  A <- as.matrix(rbind(A_top, setNames(A_btm, colnames(A_top))))
  rownames(A) <- paste(c(rep(c(state, "RoUS"), each = length(commodities)),
                         rep(c(state, "RoUS"), each = length(industries))),
                       c(rep(commodities, 2), rep(industries, 2)),
                       c(rep("Commodity", length(commodities)*2),
                         rep("Industry", length(industries)*2)),
                       sep = ".")
  colnames(A) <- rownames(A)

  logging::loginfo("Generating the L matrix ...")
  # Calculate L matrix
  I <- diag(nrow(A))
  L <- solve(I - A, tol = 1E-20)

  logging::loginfo("Calculating y (Final Demand totals) of SoI and RoUS ...")
  # Calculate Final Demand (y)
  FD_columns  <- getFinalDemandCodes("Summary")
  ita_column <- ifelse(iolevel == "Detail", "F05100", "F051")
  SoI2SoI_y   <- rowSums(ls[["SoI2SoI"]][, c(FD_columns, ita_column, "ExportResidual")])
  SoI2RoUS_y  <- rowSums(ls[["SoI2RoUS"]][, c(FD_columns, ita_column)])
  RoUS2SoI_y  <- rowSums(ls[["RoUS2SoI"]][, c(FD_columns, ita_column)])
  RoUS2RoUS_y <- rowSums(ls[["RoUS2RoUS"]][, c(FD_columns, ita_column, "ExportResidual")])
  y <- c(SoI2SoI_y + SoI2RoUS_y, RoUS2SoI_y + RoUS2RoUS_y, rep(0, length(industries)*2))
  names(y) <- rownames(L)

  logging::loginfo("Validating L*y == commodity and industry output ...")
  # Validate L * y == Output
  # Output = c(SoI_TCO, RoUS_TCO, SoI_TIO, RoUS_TIO)
  output <- c(SoI_Commodity_Output, RoUS_Commodity_Output,
              SoI_Industry_Output, RoUS_Industry_Output)
  validation <- as.data.frame(L %*% y - output)
  colnames(validation) <- "L*y-output"

  validation$rel_diff <- validation$`L*y-output`/output
  validation$Ly <- as.numeric(L %*% y)
  validation$output <- output
  validation[validation$output == 1, "rel_diff"] <- 0

  logging::loginfo("Validation complete.")
  return(list(A = A, L = L, y = y, Validation = validation))
}

# Generate validation results
GA_2r_LagaintsOutput_Validation <- validateTwoRegionLagainstOutput("Georgia", year, ioschema = 2012, iolevel = "Summary")
MN_2r_LagaintsOutput_Validation <- validateTwoRegionLagainstOutput("Minnesota", year, ioschema = 2012, iolevel = "Summary")
OR_2r_LagaintsOutput_Validation <- validateTwoRegionLagainstOutput("Oregon", year, ioschema = 2012, iolevel = "Summary")
WA_2r_LagaintsOutput_Validation <- validateTwoRegionLagainstOutput("Washington", year, ioschema = 2012, iolevel = "Summary")

# Failures
GA_failures <- formatValidationResult(GA_2r_LagaintsOutput_Validation$Validation[, "L*y-output", drop = FALSE],
                                      abs_diff = TRUE, tolerance = 1E6)[["Failure"]]
MN_failures <- formatValidationResult(MN_2r_LagaintsOutput_Validation$Validation[, "L*y-output", drop = FALSE],
                                      abs_diff = TRUE, tolerance = 1E6)[["Failure"]]
OR_failures <- formatValidationResult(OR_2r_LagaintsOutput_Validation$Validation[, "L*y-output", drop = FALSE],
                                      abs_diff = TRUE, tolerance = 1E6)[["Failure"]]
WA_failures <- formatValidationResult(WA_2r_LagaintsOutput_Validation$Validation[, "L*y-output", drop = FALSE],
                                      abs_diff = TRUE, tolerance = 1E6)[["Failure"]]
GA_failures_reldiff <- formatValidationResult(GA_2r_LagaintsOutput_Validation$Validation[, "rel_diff", drop = FALSE],
                                              abs_diff = TRUE, tolerance = 1E-2)[["Failure"]]
MN_failures_reldiff <- formatValidationResult(MN_2r_LagaintsOutput_Validation$Validation[, "rel_diff", drop = FALSE],
                                              abs_diff = TRUE, tolerance = 1E-2)[["Failure"]]
OR_failures_reldiff <- formatValidationResult(OR_2r_LagaintsOutput_Validation$Validation[, "rel_diff", drop = FALSE],
                                              abs_diff = TRUE, tolerance = 1E-2)[["Failure"]]
WA_failures_reldiff <- formatValidationResult(WA_2r_LagaintsOutput_Validation$Validation[, "rel_diff", drop = FALSE],
                                              abs_diff = TRUE, tolerance = 1E-2)[["Failure"]]
```

```{r report-results}
# GA
cat("##### 18.1 Georgia and Rest of the US\n")
cat("Absolute Difference: ")
reportValidationResult(GA_failures)
cat("Relative Difference: ")
reportValidationResult(GA_failures_reldiff)
# MN
cat("##### 18.2 Minnesota and Rest of the US\n")
cat("Absolute Difference: ")
reportValidationResult(MN_failures)
cat("Relative Difference: ")
reportValidationResult(MN_failures_reldiff)
# OR
cat("##### 18.3 Oregon and Rest of the US\n")
cat("Absolute Difference: ")
reportValidationResult(OR_failures)
cat("Relative Difference: ")
reportValidationResult(OR_failures_reldiff)
# WA
cat("##### 18.4 Washington and Rest of the US\n")
cat("Absolute Difference: ")
reportValidationResult(WA_failures)
cat("Relative Difference: ")
reportValidationResult(WA_failures_reldiff)
```
